---
title: "GeoexperimentsResearch version 1.0"
date: "`r Sys.Date()`"
author: "Google, Inc."
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
  md_document:
    toc: true
  g3doc::md_document:
    toc: true
---
```{r setup, echo = FALSE, message = FALSE}
set.seed(20170224L)
options(width=99)
```

# Geo Experiments: an introduction

## What are geo experiments?

A geo experiment is a controlled online experiment, with the aim of quantifying
the effects of treatments using geographical regions as experimental units.

In its simplest form, a geographical region is divided into a Control and a
Treatment region; the people in the Treatment region are exposed to a certain
treatment, while for those in the Control region nothing changes. The behavior
(response of some given observable variable) in both these regions is observed
and measured and the effect of the treatment is estimated using a statistical
model.

The Control and Treatment regions are formed by aggregating smaller geographical
units (called 'geos'), usually by randomization or by a matching algorithm.

### Geo experiments for estimating ad effectiveness

A typical application is to investigate the effect of increased investment in
online advertising. People who reside in the Treatment region are served ads at
a higher intensity (spend per time unit) whenever they do a web search using
prespecified keywords. Those who reside in the Control region continue to be
served ads at the usual intensity. Such an experiment typically lasts for a
few weeks.

The aim of a geo experiment in the online ad context is to estimate the return
(in monetary terms) on the money spent: more accurately the incremental return
on ad spend (iROAS), which refers to the additional return that would have not
been received without the specific additional spend in ads. The "return" is
usually the aggregate sales, either online or offline.

Similarly, we can estimate the effect of decreased spending. So the market
intervention here may be either "increased" spending, "decreased" spending, or
in general, some spend "change". In the simple one-period geo experiments, the
statistical model only needs to know which geos experienced a spend change and
which did not (i.e., which geos were in the Control regions).


## Overview of this vignette

This vignette shows briefly how to use the R package 'GeoexperimentsResearch,'

1. to represent information as data objects;
1. to analyze geo experiment data once an experiment has finished;
1. to run a preanalysis.

For details, refer to the package manual.

For further general information on geo experiments, see [1], [2], and [3].


## Attaching the package

```{r, message=FALSE, echo=TRUE}
library(GeoexperimentsResearch)
```

## Structure of geo experiment data

For the purpose of illustration, we assume that we have run a geo experiment
and wish to estimate the incremental return on ad spend.

We need to collect three pieces of information:

* The experiment periods (date ranges of the Pretest, intervention, and
  cooldown periods);
* The geo assignment;
* The observational data for both the response metric and the cost
  metric.

### Experiment periods

A geo experiment consists of several, distinct time periods: the Pretest,
Intervention, and Cooldown periods.  The latter two periods combined make up
the `Test' period.

During the Pretest period, any ad campaigns in the Treatment and Control geos
that are targeted by the experiment are in their unmodified base state.  All
geos operate with the same baseline level campaign settings (e.g., common
bidding, keyword lists, ad targeting, etc); the difference between the Control
and Treatment geos is zero in expectation.

The targeted ad campaigns are modified in the Treatment geos during the
Intervention period.

Finally, these targeted ad campaigns are reset to their original state during
the Cooldown period.  This does not always mean their effects will cease
instantly.  Incremental offline sales, for example, may continue to accrue
across subsequent days or even weeks.  Including data from the Cooldown period
in the analysis makes it possible to capture these lagged effects from the
advertising change.  This lagged impact may be substantial or not, depending on
the advertising situation; hence it can be excluded if it is obvious in the
analysis that there are no lagged effects.

By convention, we number the periods as 0 (Pretest), 1 (Intervention), 2
(Cooldown), but other numbering is allowed provided that the order of the
periods is unchanged.

This information is represented by the *ExperimentPeriods* object class.  The
start dates of each period must be specified, and finally end date of the
experiment.  This example has only a pretest period and one intervention period:
```{r message=TRUE, echo=FALSE}
  print(ExperimentPeriods(c("2015-01-05", "2015-02-16", "2015-03-15")))
```

### Geo assignment

Before the experiment is started, each of the geos is assigned either to
Control or to Treatment region (_geo group_). This mapping between a geo to the
geo group is called the _geo assignment_.

This information is represented by the *GeoAssignment* object class.
Example:
```{r message=TRUE, echo=FALSE}
data(geoassignment)
head(GeoAssignment(geoassignment))
```


### Observational data

The observational data consist of a response metric (such as sales) and cost
metric (such as cost of ad clicks). These are provided broken by date and geo,
including the Pretest, Intervention, and Cooldown periods. If the data is weekly
data, the weekly aggregate should be associated with the same day of the month,
for instance, each Sunday.

This information is represented by the *GeoTimeseries* object class.  Example: a
few rows shown from the example data set:

```{r message=TRUE, echo=FALSE}
data(salesandcost)
head(GeoTimeseries(salesandcost, metrics=c("sales", "cost")))
```

# Analyzing geo experiment data using the GBR and TBR methods

## Reading in observational data

Load the sample data set.

```{r message=FALSE, echo=FALSE}
data(salesandcost)
```
This is a plain *data.frame*, with the following columns:

```{r message=FALSE, echo=TRUE}
head(salesandcost)
```

This data frame has a date and a geo column, and two _metrics_, the sales and
the cost of ad clicks, numeric values that are associated to each geo and date.

Next, we convert this data frame into a *GeoTimeseries* object and the integrity
of the time series is automatically checked. We need to specify which columns
are to be treated as metrics. This helps the class methods do certain
operations automatically, such as aggregation over geos and time.

```{r message=FALSE, echo=TRUE}
obj.gts <- GeoTimeseries(salesandcost, metrics=c("sales", "cost"))
```

No errors occurred, so the overall structure of the data seems to be fine.  The
resulting object inherits from *data.frame*, with the same columns, augmented
with some extra columns:

```{r message=FALSE, echo=TRUE}
head(obj.gts)
```
> The 'date' column must be in either 'Date', factor, or character format and is
> always coerced to Date. If the date format differs from 'yyyy-mm-dd', it is
> necessary to specify it as argument 'date.format'.

> The column 'geo' is of type character even though some geo IDs (such as DMAs)
> are represented as integers. Using character format, however, the structure
> of *GeoTimeseries* is also compatible with non-integer geos such as postal
> codes and administrative regions without remapping them to numbers.

> There is no checking of whether any of the metrics are negative.

There are some extra columns provided for convenience:

* _.weekday_ column denotes the day of the week (1=Monday, 7=Sunday);
* _.weeknum_ indicates the number of the week within a year;
* _.weekindex_ indicates a unique week number.

> The data frame can have any number of other columns, although the built-in
> methods recognize only 'date', 'geo', '.weekday', '.weeknum', and
> '.weekindex', and those registered as metrics.


### Exploratory data analysis

To quickly investigate the distribution of the metrics across weeks, we can use
the *aggregate* method as follows:

```{r message=FALSE, echo=TRUE}
aggregate(obj.gts, by='.weekindex')
```
We can see that normally the ad campaigns were turned off, and starting from
week 6, the ad spend increased until it was turned off again on week 14.

To plot the time series, use the plot method:

```{r gts-plot, message=FALSE, echo=TRUE, eval=TRUE}
 plot(obj.gts)
```

To hide the legend, add `legend=FALSE`. To plot the time series on log scale,
add `log.scale=TRUE`.  For more information of the method, type
`?plot.GeoTimeseries` at the R prompt.

## Experiment Periods

We specify the start of the Pretest period, the start of the test period, and
the end of the experiment.  If there is a Cooldown period after the actual
market intervention, it must be included as a separate period (four dates in
total).

```{r message=FALSE, echo=TRUE}
obj.per <- ExperimentPeriods(c("2015-01-05", "2015-02-16", "2015-03-15"))
obj.per
```
To introduce a cooldown period, we would specify one more date.

To learn more about the function, type `?ExperimentPeriods` at the R prompt.


## Geo Assignment

We'll use the built-in sample geo assignment:

```{r message=FALSE, echo=TRUE}
data(geoassignment)
head(geoassignment)
```

From this data frame we create a *GeoAssignment* object and automatically
verify its integrity:

```{r message=FALSE, echo=TRUE}
obj.ga <- GeoAssignment(geoassignment)
head(obj.ga)
```

## Combining all information about the experiment into one object

The class *GeoExperimentData* combines these three pieces of information (geo
time series, periods, geo assignment) into one object:

```{r message=FALSE, echo=TRUE}
  obj <- GeoExperimentData(obj.gts,
                           periods=obj.per,
                           geo.assignment=obj.ga)
  head(obj)
```

The column *period* contains the indicator for the experiment periods: 0 =
Pretest, 1 = test (Intervention). 'NA' marks a date that is outside of the
designated experiment periods.

The column *geo.group* contains the geo group ID for each of the geos.

The column *assignment* is not used in this version of the R package. It is set
to *NA* by default. It can be ignored.


### Exploratory data analysis

To check how the revenue and cost metrics are distributed across periods and
groups, we make use of the aggregate method again:
```{r message=FALSE, echo=TRUE}
aggregate(obj, by=c('period', 'geo.group'))
```


## Geo-Based Regression (GBR) Analysis

The object ('obj') that we constructed contains now all information for
applying a geo experiment analysis methodology.

To perform a GBR (geo-based regression) analysis, apply method
*DoGBRROASAnalysis*, specifying which of the metrics is the response and which
represents the cost, along with the experiment periods and group numbers.

```{r message=FALSE, echo=TRUE}
  result <- DoGBRROASAnalysis(obj, response='sales', cost='cost',
                              pretest.period=0,
                              intervention.period=1,
                              cooldown.period=NULL,
                              control.group=1,
                              treatment.group=2)
  result
```

Note that in this particular case, there is no `cooldown.period`, hence it is
set to `NULL`.  If there was one, we would specify the period number (for
example, `cooldown.period=2`).

The resulting object (a *GBRROASAnalysisFit* object) contains the model fit:
when printed, it shows its summary, which defaults to 90 percent credible
intervals. To recalculate the interval with a different credibility level, we
can specify this in the function call:

```{r message=FALSE, echo=TRUE}
  summary(result, level=0.95, interval.type="two-sided")
```

To obtain the posterior probability that the true iROAS is larger than some
threshold, say 3.0, we use the `summary` method as follows:
```{r message=FALSE, echo=TRUE}
  summary(result, threshold=3.0)
```
The default threshold is 0.


## Time-Based Regression (TBR) ROAS Analysis

The GeoExperimentData object can also be used for performing a TBR analysis [3],
applying method *DoTBRROASAnalysis*, specifying which of the metrics is the
response and which represents the cost, along with the experiment period and
group numbers. The model ID is also required; currently the only available
model is 'tbr1', as described in [3].

```{r message=FALSE, echo=TRUE}
  obj.tbr.roas <- DoTBRROASAnalysis(obj, response='sales', cost='cost',
                                    model='tbr1',
                                    pretest.period=0,
                                    intervention.period=1,
                                    cooldown.period=NULL,
                                    control.group=1,
                                    treatment.group=2)
  obj.tbr.roas
```

The resulting object (a *TBRROASAnalysisFit* object) contains the model fit:
when printed, it shows its summary, which defaults to 90 percent one-sided
credible intervals. Similarly to what we did with a *GBRROASAnalysisFit* object
we can recalculate the credible interval, and the probability of exceeding a
given threshold like so:

```{r message=FALSE, echo=TRUE}
  summary(obj.tbr.roas, level=0.95, interval.type="two-sided")
```

```{r message=FALSE, echo=TRUE}
  summary(obj.tbr.roas, threshold=3.0)
```

The *plot* method shows the evolution of the iROAS estimate across the Test
period:

```{r tbr-roas-plot, message=FALSE, echo=TRUE, eval=TRUE}
 plot(obj.tbr.roas)
```

For more information on the method, type
`?plot.TBRROASAnalysisFit` at the R prompt.



## Time-Based Regression (TBR) Causal Effect Analysis

Unlike the TBR ROAS Analysis, which estimates the ratio of the incremental
response and incremental cost, the TBR Causal Effect Analysis applies only to
one single variable, such as revenue.

```{r message=FALSE, echo=TRUE}
  obj.tbr <- DoTBRAnalysis(obj, response='sales',
                           model='tbr1',
                           pretest.period=0,
                           intervention.period=1,
                           cooldown.period=NULL,
                           control.group=1,
                           treatment.group=2)
```

The resulting object (a *TBRAnalysisFitTbr1* object) contains the model fit for
each time point, which can be seen when printed. To show the summary of the
effect, we use the *summary* method:

```{r message=FALSE, echo=TRUE}
  summary(obj.tbr)
```
which defaults to the 90% one-sided interval.

The *plot* method illustrates the results of the analysis.
```{r tbr-plot, message=FALSE, echo=TRUE, eval=TRUE}
 plot(obj.tbr)
```

For more information on the method, type
`?plot.TBRAnalysisFitTbr1` at the R prompt.

# Preanalysis

Before running an experiment, we need to understand how the design parameters
affect the uncertainty of the iROAS estimate. One of the most important
parameters is the ad spend change, which affects the estimate uncertainty
directly: doubling the ad spend halves the width of the 2-sided confidence
interval (in terms of one-sided intervals, this is the distance from the lower
bound and the point estimate). We refer to this confidence interval half-width
by _precision_ (which gets better as the confidence interval gets shorter).

The function *DoROASPreanalysis* predicts the precision of the iROAS estimate
based on historical data provided. It simulates experiments (by resampling) with
given period lengths and records the precision from each simulated
experiment. We can then use the *summary* method to compute the precision given
an ad spend change, or find the ad spend change associated with a given
precision.

For each simulated geo experiment data set, ROAS and its precision is estimated.
The process yields a distribution of the these estimates of precision.  The
summary method takes the empirical median as the point estimate.  If the data
set does not have strong seasonalities, the variation of this estimate should be
fairly small.

The process runs as follows:

1. Assign geos to treatment groups.
1. Run preanalysis to predict the precision.

 

## A randomized geo assignment

Randomized geo assignments can be done using `GeoStrata' objects.  This object
includes a mapping from each geo to a _stratum_ (or block), so a stratified
randomization can be performed. This can be generated automatically using the
*ExtractGeoStrata* function:
```{r message=FALSE, echo=TRUE}
obj.geo.strata <- ExtractGeoStrata(obj.gts, volume="sales", n.groups=2)
head(obj.geo.strata)
```
The argument 'volume' specifies the name of the metric that is used for
stratification: the geos are sorted by their volume and divided into strata of 2
each.

To generate a randomized geo assignment, we use the `Randomize' method:
```{r message=FALSE, echo=TRUE}
obj.geo.assignment <- Randomize(obj.geo.strata)
head(obj.geo.assignment)
```

## Predicting the precision

We pass this object to the method *DoGBRPreanalysis* along with the
GeoTimeseries, the length of the Pretest, Intervention, and Cooldown periods,
and specify a metric:
```{r message=FALSE, echo=TRUE, eval=TRUE}
obj.pre <-  DoROASPreanalysis(obj.gts, response="sales",
                              geos=obj.geo.assignment,
                              prop.to="sales",
                              period.lengths=c(42, 21, 7))
```
The resulting object 'obj.pre' is of class *ROASPreanalysisFit*, which only
contains the raw simulated numbers. To compute the required spend for precision
+/- 1.0, we call the *summary* method:

```{r message=FALSE, echo=TRUE}
  results <- summary(obj.pre,
               level=0.90,
               type="one-sided",
               precision=1.0)
  print(results)
```

This function takes the median of the simulated precisions as the point
estimate.

For convenience, applying the 'print' method on the "GBRPreanalysisFit" prints
the default summary (one-sided confidence interval at level 0.90, precision
1.0):

```{r message=FALSE, echo=TRUE}
  print(obj.pre)
```

The function can be also used to predict the precision given a (total) spend
change over the test period:

```{r message=FALSE, echo=TRUE}
  results2 <- summary(obj.pre,
                level=0.90,
                type="one-sided",
                cost=10000)
  print(results2)
```

The results apply to the given geo assignment only; for a different geo
assignment, the results are likely to be different.


# References

[1] Kerman, J., Vaver, J. and Koehler, J. (2011)
    [Estimating causal effects using geo experiments](http://www.unofficialgoogledatascience.com/2016/06/estimating-causal-effects-using-geo.html)

[2] Vaver, J. and Koehler, J. (2011)
    [Measuring Ad Effectiveness Using Geo Experiments](http://static.googleusercontent.com/media/research.google.com/en//pubs/archive/38355.pdf)

[3] Kerman, J. and Wang, P., and Vaver, J. (2017)
    [Estimating Ad Effectiveness using Geo Experiments in a Time-Based Regression Framework](https://research.google.com/pubs/pub45950.html)

# Disclaimer
This software is not an official Google product. For research purposes only. Copyright 2017 Google, Inc.

